

Included in R and various R packages are some basic datasets that are useful for testing functions and learning about R features and functions. One such dataset is **cars**. To learn about this dataset you can simple use the `{r} help` function

```{r}
help(cars)
```
Use the `{r} str()` and `{r} summary()' functiosn to see basic summary statistics about this dataset
```{r}
str(cars)
summary(cars)
```
To visualize this data you can generate a simple plot with the `{r} plot()` function
```{r}
plot(cars)
```
You can also embed plots, for example: # JTL, line by line, got an error; plus not sure how useful? what's point?
```{r, echo=FALSE}
plot(cars)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.'

## Other Useful Features and Fucntions: Sorting, Subsetting, Sampling
##JTL: seems like some of this could be combined with stuff above
**Sorting**
We can use another dataset (mtcars) to practice sorting (ordering) data. Learn about mtcars via `{r} help(mtcars)`

sort by mpg
```{r}
newdata <- mtcars[order(mtcars$mpg),]
```
sort by mpg and cyl # JTL: not sure how effect the cyl sort is
```{r}
newdata <- mtcars[order(mtcars$mpg, mtcars$cyl),]
```
sort by mpg (ascending) and cyl (descending) #JTL: same as above?
```{r}
newdata <- mtcars[order(mtcars$mpg, - mtcars$cyl),]
```
Now, Let's make a new vector of data
```{r}
z <- c(1.5, 1/6, 1/3)
```
If we only want to view  the first two decimal places of z
```{r}
round(z,2)
```
Now, we can reverse the order of the elements in z
```{r}
rev(z)
```
And we can order z from smallest to largest
```{r}
sort(z)
```
We can also identify the ordering of z #JTL: with respect to what?
```{r}
order(z)
```
i.e., the 2nd number is the min and the 1st number is the max

Additionally, we can idenify the maximum values this way:
```{r}
max(z)
```
**Subsetting**
Let's create a original object vector, x:
```{r}
x <- c(3, 4, 7)
x
```
Now, let's subset this vector and keep only the first three values
```{r}
x[-3]
```
Now, let's subset this vector and keep only the velues greater than or equal to 5
```{r}
x[x >= 5]
```
Notice that we did this using a logic statement `{r} >=`. Here is a list of other logica operators that you might find useful:

|Logic Operator|Meaning| # confusing to start using new symbols "|"?
|! x | Is Not "x"|
|x & y| "x" and "y" (element by element) |
|x && y| "x" and "y" (across all elements)|
|x `|` y | "x" or "y" (element by element)|
|x `||` y | "x" or "y" (across all elments)|

You can learn more about this commands (`{r} help(Logic, package=base))


**Sampling**








First, let's create a sequence of numbers
```{r}
seq(1,3,length=5)

# Create the same sequence in a slightly different way:
seq(1,3,by=0.5)

# Create another sequence by going from 3 to 1:
seq(3,1,by= -0.5)
```

To randomly sample from an existing vector:
```{r}
sample(x,10,replace=T)
```

Or to randomly sample from a sequence of numbers from 1 to 500:
```{r}
sample(1:500,10,replace=F)
```




